Comments on the revised version

### 2018-03-07 paper/paper.tex
updates

Providing the computational infrastuctures needed to solve actual complex
problems arising in the various fields of modern society (including climate
change, health, green energy or security) is a strategic challenge. The main
pillar to adress this problem is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity
leads to sophisticated specific Exascale platforms, and the objective of the
community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly,
system management software still has a long way to go. The existing job and
resource management software allow to run tens of thousands of jobs on hundreds
of thousands of cores. Such software are based on robust policies with
positive properties such as relative simplicity and prevention of job
starvation. However, there is still room for improvement. Here, we study how to
tune the ordering of the submission queues in the frame of the classical
so-called EASY\footnote{Extensible Argonne Scheduling sYstem
(EASY)}-Backfilling family of heuristics using two new online approaches. One
approach is based on a simulator and the second uses a multi-armed bandit
algorithm.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. In this work, we focus on the
average waiting time of the jobs. Each method aims at choosing the best queue
reordering policy in a fixed search set of semantically diverse options.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in~\cite{feitresampling}, an approach that attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, as illustrated in~\cite{jsspp17},
it is reasonable
to assume that the system is in a stationary state.~\cite{jsspp17} and assume for these experiments that the system is in a
stationary state. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. In our experiments, we use stationary resampled data, 
and therefore no decay is necessary: $\lambda$ here is set to $1$.

@@

  \label{tab:full}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & LCFS  & LPF            & SPF   & LQF            & SQF   & SEXP  & LEXP           & LRF            & SRF   & LAF            & SAF   \\
    \hline
    KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%          & +15\% \\
    CTC-SP2      & -40\% & -19\%          & -14\% & \textbf{-47\%} & +15\% & -36\% & -19\%          & -38\%          & -12\% & -44\%          & +22\% \\
    SDSC-SP2     & -8\%  & \textbf{-15\%} & +4\%  & -8\%           & +4\%  & -3\%  & \textbf{-15\%} & -3\%           & -10\% & -11\%          & +18\% \\
    SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%          & +5\%  \\
    CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%          & -17\% \\
    Unilu-Gaia   & -58\% & -33\%          & -10\% & \textbf{-62\%} & +16\% & -49\% & -27\%          & -56\%          & -16\% & -62\%          & +20\% \\
    \hline
  \end{tabular}
\end{table*}

@@

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
    \hline
    Trace        & Best(hindsight) & Random &       & Full           &                & Noisy &       & Bandit & \\
    \hline                
    $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & -\textbf{11}\% & -12\% & -12\% & -7\%   & -10\% \\
    CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -13\% & -13\% & -5\%   & -8\% \\
    SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -53\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

@@

There are two directions to extend this work. First, the question of
the existence of a more efficient selection strategy for bandit feedback for
these systems is not addressed by this work. Is a simulator really necessary?
%@Denis: a essayer.. 1 mois de travail pour tester?
Second, one can't help but wonder what performance could be achieved by extending
the search space to a wider class of policies. This means studying learning performance using some
function approximator.
%@Denis: c'est ce que font vasily et salah de deux manieres differentes.
%TODO SPLIT IN TWO, MENTION FACTOR ANALYSIS.
As both these directions are data-bound, they will require careful application of
trace resampling techniques to be successful.

>>>>

Providing the computational infrastuctures needed to solve actual complex
problems arising in the various fields of modern society (including climate
change, health, green energy or security) is a strategic challenge. The main
pillar to adress this problem is to build extreme-scale parallel and distributed
platforms. The never-ending race for more computing power and storage capacity
leads to sophisticated specific Exascale platforms, and the objective of the
community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly,
system management software still has a long way to go. The existing job and
resource management software allow to run tens of thousands of jobs on hundreds
of thousands of cores. Such software are based on robust policies with
positive properties such as relative simplicity and prevention of job
starvation. However, there is still room for improvement. Here, we study how to
tune the ordering of the submission queues in the frame of the classical
so-called EASY\footnote{Extensible Argonne Scheduling sYstem
(EASY)}-Backfilling family of heuristics using two online approaches. One
approach is based on a simulator and the second uses a multi-armed bandit
algorithm.

@@

Our main contribution is to investigate two methods for learning a
good scheduling policy on a per-system basis. In this work, we focus on the
average waiting time of the jobs. Each method aims at choosing the best queue
reordering policy in a fixed search set of semantically diverse options.

@@

Comparing algorithms in batch scheduling is made especially difficult by the
fluctuations in commonly used metrics~\cite{feitelson1998metrics}. The variability in
performance is known to outrank the available sample sizes: \textit{It is not
enough to simply replay an algorithm using a workload trace}. Indeed, as
presented in Figure~\ref{fig:variability}, the objective function we are using
in this work is subject to high variability. In our experience, randomly
shuffling months in a trace at random is enough to generate contradictory
conclusions about which algorithm is better. This is a known
problem~\cite{flurries} that motivated decades of research in workload
modeling~\cite{feitbook}. Workload models have drawbacks in that they usually
lose the details of the workloads. This is problematic due to the non-linear
aspect of the job scheduling process. This is discussed
in~\cite{feitresampling}, an approach that attempts to solve this problem by
resampling data directly from the original traces. In this work, we use a
simple implementation of these ideas. More precisely, we want to sample the job
submission process, which is achieved in the following manner: (a) For every
week in a resampled trace, we iterate over every existing user in the original
trace; (b) For each of these users, we choose a week of the original trace
uniformly at random; (c) The jobs from this user in this original week are
chosen to be present in the week being resampled. This simple mechanism enables
us to preserve the dependency within a week. Although the system state may not
be independent from one week to an other, as illustrated in~\cite{jsspp17},
it is reasonable
to assume that the system is in a stationary state. Indeed,~\cite{jsspp17} performs
train/test experiments using the same policy space as we use here. We do provide guidelines as how to incorporate tracking
into our algorithms in Section~\ref{sec:online} by introducing a decaying
discount factor $\lambda$. In our experiments, we use stationary resampled data,
and therefore no decay is necessary: $\lambda$ here is set to $1$.

@@

  \label{tab:full}
  \begin{tabular}{@{}lrrrrrrrrrrr@{}}
    \hline
    Trace        & LCFS  & SPF            & LPF   & SQF            & LQF   & LEXP  & SEXP           & SRF            & LRF   & SAF            & LAF   \\
    \hline
    KTH-SP2      & -13\% & \textbf{-16\%} & +5\%  & \textbf{-16\%} & +3\%  & -8\%  & -15\%          & -8\%           & -13\% & -12\%          & +15\% \\
    CTC-SP2      & -40\% & -19\%          & -14\% & \textbf{-47\%} & +15\% & -36\% & -19\%          & -38\%          & -12\% & -44\%          & +22\% \\
    SDSC-SP2     & -8\%  & \textbf{-15\%} & +4\%  & -8\%           & +4\%  & -3\%  & \textbf{-15\%} & -3\%           & -10\% & -11\%          & +18\% \\
    SDSC-BLUE    & -26\% & -29\%          & 0\%   & -29\%          & +17\% & -17\% & -26\%          & -15\%          & -15\% & \textbf{-32\%} & +23\% \\
    ANL-Intrepid & -25\% & -9\%           & -3\%  & \textbf{-28\%} & -2\%  & -21\% & -24\%          & -21\%          & -10\% & -19\%          & +5\%  \\
    CEA-Curie    & -46\% & -23\%          & -45\% & \textbf{-57\%} & -19\% & -51\% & -23\%          & \textbf{-57\%} & -24\% & -40\%          & -17\% \\
    Unilu-Gaia   & -58\% & -33\%          & -10\% & \textbf{-62\%} & +16\% & -49\% & -27\%          & -56\%          & -16\% & -62\%          & +20\% \\
    \hline
  \end{tabular}
\end{table*}

@@

\begin{table*}[]
  \centering
  \ra{1.3}
  \caption{Average Cumulative waiting time improvement of the policies used with respect to EASY(FCFS).}
  \label{tab:strat}
  \begin{tabular}{@{}lrrrrrrrrrrrr@{}}
    \hline
    Trace        & Best(hindsight) & & Random       &            &Full                & & Noisy        & & Bandit  \\
    \hline
    $\Delta$     &        & week   & day   & week           & day            & week  & day   & week   & day \\
    KTH-SP2      & -16\%  & -6\%   & -8\%  & \textbf{-12}\% & 11\% & -\textbf{-12}\% & -\textbf{-12}\% & -7\%   & -10\% \\
    CTC-SP2      & -47\%  & -20\%  & -23\% & \textbf{-44}\% & \textbf{-44}\% & -44\% & -44\% & -26\%  & -31\% \\
    SDSC-SP2     & -15\%  & -4\%   & -7\%  & \textbf{-13}\% & \textbf{-13}\% & -\textbf{-13}\% & -\textbf{-13}\% & -5\%   & -8\% \\
    SDSC-BLUE    & -32\%  & -12\%  & -14\% & \textbf{-31}\% & \textbf{-31}\% & -30\% & -29\% & -12\%  & -17\% \\
    ANL-Intrepid & -28\%  & -13\%  & -20\% & -22\%          & \textbf{-26}\% & -19\% & -24\% & -13\%  & -20\% \\
    CEA-Curie    & -57\%  & -35\%  & -42\% & \textbf{-53}\% & -47\%          & -\textbf{-53}\% & -48\% & -36\%  & -46\% \\
    Unilu-Gaia   & -62\%  & -29\%  & -31\% & -59\%          & \textbf{-60}\% & -58\% & -58\% & -33\%  & -34\% \\
    \hline
  \end{tabular}
\end{table*}

@@

There are three directions to extend this work. First, the question of the
existence of a more efficient selection strategy for bandit feedback for these
systems is not addressed by this work. Is a simulator really necessary?
Second, one can't help but wonder what performance could be achieved by
extending the search space to a wider class of policies. This means studying
learning performance using some function approximator.

As both these directions are data-bound, they will require careful application of
trace resampling techniques to be successful.

### 2018-03-07 paper/revision2.tex
updates

Let us first sincerely thank the reviewers for their relevant and in-depth
remarks. Their comments helped us to clarify the focus of this work, which
lead to restructure the evaluation procedure to some extent. As part of these
modifications, we changed the title to: "Online Tuning of EASY-Backfilling
using Queue Reordering Policies". It better reflects the main goal, which is
not to advocate bandit strategies, but rather to study their feasibility, and
to show where the tradeoff is compared to using simulations for tuning the
algorithms. We hope this revised version may now be considered for publication. 
We detail our answers to rewiewer's comments (bold).

@@

\textbf{Last paragraph of 1.2 is the third repetition of this.  It is redundant, despite the fact that some reviewer didn't get
it.
}
You are right. We rephrased this in a more neutral way.

\textbf{In sect 5.1 you explain that you set lambda=1 because due to the resampling the distributions are stationary.  That is
correct.  But it would have been good to also conduct simulations on the original traces, which are probably not
stationary, and therefore there is more to gain from switching between policies.
Also, instead of starting the simulation of each period from an empty system, why not start from the situation at the
end of the previous period?
}
Regarding your first point, good remark, we based this experimental design around previous results
from "Tuning Backfilling Queues"(JSSPP17) that show some degree of stationarity. We indeed
expect that non-stationary policy adaptation may allow more improvements (while
being harder to achieve). We made this clearer in the text.

Your second point is also relevant, yes that should perhaps slightly improve the results.
But we can see that the 'full feedback' simulated policy here is almost as good as the best
policy in the set, so we did not worry about this. In a non-stationary case, it may indeed
be necessary to properly take care of this.

\textbf{Typo at end of 6.1.3: sentence with [24] appears twice.
}
Right, we fixed this.

@@

\textbf{The authors present a study based on a selection of several strategies and parameters. They execute several simulations
and show that using a feedback mechanism improvements over the standard approach without feedback are possible. This
results is not surprising considering previous results some of which are referenced within the manuscript. It must be
noted that the authors do not suggest a specific new scheduling algorithm and show that this algorithm results in
significant improvements over existing methods.
}
We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
results in significant improvements over existing methods.
Thank you for your in-depth remarks once again. We integrated these into the minor revision.

@@

First of all, we corrected a crucial typographical error in the text. The S and
L were inverted in both Table 3 and Figure 4. This resolves a blatant
consistency issue with other publications, most crucially "Tuning Backfilling
Queues"(JSSPP17).

@@

However, a good estimation of this quantity requires using more traces, and perhaps factoring
in the topology model of the machines, which should increase the heterogeneity between platforms
and therefore increase the dependence of the simulation performance on the trace/platform.
Such a study would be much more tedious.
Here, we prove the feasibility of the approach. The need for adaptivity is actually understated
by our simulation approach.

@@

>>>>

Let us first sincerely thank the reviewers for their relevant and in-depth
remarks once again. Their comments helped us further the discussion of the
results, and find a typographical error that passed our scrutiny.

We detail our answers to rewiewer's comments (bold).

@@

\textbf{Last paragraph of 1.2 is the third repetition of this.  It is redundant, despite the fact that some reviewer didn't get
it.
}

You are right. We rephrased this in a more neutral way.

\textbf{In sect 5.1 you explain that you set lambda=1 because due to the resampling the distributions are stationary.  That is
correct.  But it would have been good to also conduct simulations on the original traces, which are probably not
stationary, and therefore there is more to gain from switching between policies.
Also, instead of starting the simulation of each period from an empty system, why not start from the situation at the
end of the previous period?
}

Regarding your first point, good remark, we based this experimental design
around previous results from "Tuning Backfilling Queues"(JSSPP17) that show
some degree of stationarity using a train/test procedure. We made this clearer
in the text.
We indeed expect that non-stationary policy adaptation may allow more
improvements (while simultaneously being harder to achieve).

Your second point is also relevant, yes that should improve the results.
But we can see that the 'full feedback' simulated policy here is almost as good as the best
policy in the set, so we did not worry about this. In a non-stationary case, it may indeed
be necessary to properly take care of this.

\textbf{Typo at end of 6.1.3: sentence with [24] appears twice.
}

Right, we fixed this.

@@

\textbf{The authors present a study based on a selection of several strategies and parameters. They execute several simulations
and show that using a feedback mechanism improvements over the standard approach without feedback are possible. This
results is not surprising considering previous results some of which are referenced within the manuscript. It must be
noted that the authors do not suggest a specific new scheduling algorithm and show that this algorithm results in
significant improvements over existing methods.
}

We regret that the detailed answer of the previous rebuttal did not convince you. We disagree with this statement. We indeed
provide a new scheduling algorithm, which adapts to the situation by selecting the correct reordering policy. This algorithm
results in significant improvements over existing methods.
We thank you for your in-depth remarks once again. We integrated these comments into the minor revision.

@@

Right, this discussion is warranted. We added paragraph X.

First of all, we corrected a crucial typographical error in the text that slipped
our scrutiny at the time of running the build chain for the publication. The S and
L were inverted in both Table 3 and Figure 4. This resolves a blatant
consistency issue with other publications, most crucially "Tuning Backfilling
Queues"(JSSPP17).

@@

However, a good estimation of this quantity requires using more traces, and
perhaps factoring in the topology model of the machines, which should increase
the heterogeneity between platforms and therefore increase the dependence of
the simulation performance on the trace/platform. Such a study would be much
more tedious. Here, we prove the feasibility of the approach. The need for
adaptivity is actually understated by our simulation approach, but in all
honesty we don't believe to be able to measure this rigorously.

@@

Right, we need to make this clearer. We added this discussion to paragraph X.

In both these cases, the scheduling decisions do not have an overhead, as the
policy is fixed during the period (one day, or week).

The bandit strategy has a negligible computational overhead at period end. Its two
computational operations are maintaining a sequential average and generating a
pseudo-random number.

The simulation-based strategy requires to simulate the system K times every
period end, where K is the number of alternative policies considered (here,
K=12). Using our simulator, This takes a few seconds. Using a more precise
simulator with topological modeling, one can argue that the overhead is still
manageable since these simulations can be easily done in parallel. The most
important constraint is that these simulations should not take too much time
compared to the period size, as to not delay the policy switch decision too
much.
